---
title: "Datafest 2023"
author: "Thai Huynh"
date: "2023-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dbplyr)
library(caret)
library(tree)
library(ggplot2)
library(MASS)
library(knitr)
library(elasticnet)
library(glmnet)
library(psych)
library(randomForest)
```


### The client information data set

```{r}
client <- read.csv("/Users/anh.vo/Downloads/datafest2023/data/datasetABA.csv") 
```


```{r}
drop <- c("CategoryUno", "SubcategoryUno", "X")
client <-  client[,!(names(client) %in% drop)]
glimpse(client)
```



### Variable appendix

AnnualIncome : Estimated annual income for the client's household

AllowedIncome: State's calculated allowed income to determine eligibility based on the client's annual income.

AskedByClientUno: Unique identifier for the client who submitted the question

ClosedOnUtc : UTC date/time stamp indicating when the client's question was closed

LegalDeadline : Court date or other legal deadline related to the client question

AllowedAssests: State's limit for the total allowed amount for all accounts (checking, savings, etc.)

BaseIncomeLimit: Base amount used for determining a client's eligibility based on income

PerHouseholdMemberIncomeLimit : Additional amount allowed per household member used for determining a client's eligibility based on income

IncomeMultiplier: Income multiplier used for determining client's eligibility based on income


```{r}
client1 <- client %>%
  mutate(AnnualIncome=as.numeric(AnnualIncome),
            AllowedIncome=as.numeric(AllowedIncome),
            CheckingBalance=as.numeric(CheckingBalance),
            SavingsBalance=as.numeric(SavingsBalance),
            InvestmentsBalance=as.numeric(InvestmentsBalance),
            Age=as.numeric(Age),
            NumberInHousehold=as.numeric(NumberInHousehold))
```



```{r}
glimpse(client1)
```


```{r}
drop <- c("PostalCode")
client1 <-  client1[,!(names(client1) %in% drop)]
```


### Model

```{r}
library(mosaic)
# select numeric variables
df <- dplyr::select_if(client1, is.numeric)

# calulate the correlations
r <- cor(df, use="complete.obs")
round(r,2)
```

```{}
test <- lm(Category ~ ., data = client)
summary(test)
```



AllowedIncome - NumberInHousehold

AllowedAssets - IncomeMultiplier



```{r}
drop <- c("CheckingBalance", "SavingsBalance", "InvestmentsBalance")
client2 <-  client1[,!(names(client1) %in% drop)]
```


**Predict the LSC category based on the client's demographic** 

Variables of interest: Category, StateAbbr, EthnicIdentity, Age, Gender, MaritalStatus, Veteran, Imprisoned, NumberInHousehold, AnnualIncome



```{r}
client3 <- client2 %>% replace(.=="NULL", "I'd rather not answer") # replace NULL
```


```{r}
client3 <- client3 %>% replace(.=="NA", "I'd rather not answer") # replace NA
```


```{r}
client3 <- client3 %>% replace(.=="", "I'd rather not answer") # Replace Blank
```


```{r}
client3 <- client3 %>% 
  mutate(TakenByAttorneyUno = ifelse(TakenByAttorneyUno == "I'd rather not answer", "None", TakenByAttorneyUno),
         ClosedByAttorneyUno = ifelse(ClosedByAttorneyUno == "I'd rather not answer", "None", ClosedByAttorneyUno))
```


```{r}
client3 <- client3 %>%
  mutate(GetReplied = ifelse(TakenByAttorneyUno=="None", "No", "Yes"))
```


```{r}
client3 <- client3 %>% drop_na()
```

------------------------------------------------------------------


```{r}
glimpse(client3)
```

```{r}
client3 <- client3 %>%
  mutate(AllowedAssets = as.character(AllowedAssets),
         BaseIncomeLimit = as.character(BaseIncomeLimit),
         PerHouseholdMemberIncomeLimit = as.character(PerHouseholdMemberIncomeLimit),
         IncomeMultiplier = as.character(IncomeMultiplier)) %>%
    dplyr::select(-Conversation, -ClosedOnUtc, -AskedByClientUno, -TakenByAttorneyUno, -ClosedByAttorneyUno, -LegalDeadline)
```


```{r}
# Get a random sample from the large data set
client.sample1 <- client3[sample(nrow(client3), size = nrow(client3)*0.01), ]
```


### Classification Tree

```{r}
set.seed(41523)
categoryTree <- train(Category ~ .,
                   data = client.sample1,
                   method="rpart", # rpart2 if using maxdepth
                   trControl=trainControl("CV",5)) 
    
categoryTree
```


```{r}
plot(categoryTree$finalModel, margin=0.015)
text(categoryTree$finalModel, cex=0.45)
```


### LASSO

```{r}
# Train the LASSO model
category_lasso <- train( Category ~ StateAbbr + EthnicIdentity + Age + Gender + MaritalStatus + Veteran + NumberInHousehold + AnnualIncome + GetReplied,
                      data=client.sample1,
                      method="glmnet", # use glmnet package
                      trControl=trainControl(method="cv", 5), # 5-fold cross-validation
                      preProcess=c("center", "scale"),
                      tuneGrid = expand.grid(alpha=1, lambda=seq(0, 0.1, by=0.01)))
# The model output
category_lasso
```


```{r}
# Best model coefficients
coef(category_lasso$finalModel, category_lasso$bestTune$lambda)
```



### Ridge Regression


```{r}
# Train the Ridge Regression model
category_ridge <- train(Category ~ StateAbbr + EthnicIdentity + Age + Gender + MaritalStatus + Veteran + NumberInHousehold + AnnualIncome + GetReplied,
                      data=client.sample1,
                      method="glmnet", # use glmnet package
                      trControl=trainControl(method="cv", 5), # 5-fold cross-validation
                      preProcess=c("center", "scale"),
                      tuneGrid = expand.grid(alpha=0, lambda=seq(0, 0.1, by=0.01)))
# The model output
category_ridge
```


```{r}
# Best model coefficients
coef(category_ridge$finalModel, category_ridge$bestTune$lambda)
```



### Enet

```{r}
# Train the elastic net model
category_enet <- train(Category ~ StateAbbr + EthnicIdentity + Age + Gender + MaritalStatus + Veteran + NumberInHousehold + AnnualIncome + GetReplied,
                   data=client.sample1,
                   method = "glmnet",
                   trControl = trainControl(method="cv", 5), # 5-fold cross-validation
                   preProcess = c("center", "scale"),
                   tuneGrid = expand.grid(alpha=seq(0, 1, by=0.1),
                                          lambda=seq(0, 0.1, by=0.01)))
# Model output
category_enet              
```

```{r}
# Best model coefficients
coefficients(category_enet$finalModel, category_enet$bestTune$lambda)
```


